
import lib
```{r}
library(readxl)
library(tidyverse)
library(janitor)
library(here)
```

loading data
```{r}
ship_data <- read_excel(here("raw_data/seabirds.xls"), sheet = 1)

```

```{r}
bird_data <- read_excel(here("raw_data/seabirds.xls"),
                        sheet = 2,
                        guess_max = 49019)

```


Cleaning colnames ship_data using janitor
```{r}
ship_data_clean_names <- clean_names(ship_data)
```
Cleaning colnames bird_data

```{r}
bird_data_clean_names <- clean_names(bird_data)
colnames(bird_data_clean_names)[3] <- "species_common_name"
colnames(bird_data_clean_names)[4] <- "species_scientific_name"

```


Cleaning the species_common_name from the column.
Getting rid of sensu lato

```{r}
bird_data_common_name_clean <- bird_data_clean_names %>% 
  separate(species_common_name, "species_common_name", "sensu" ,remove = TRUE)
```
Geting rid of CAPITAL LETTERS
```{r}
patern <- "[A-Z][A-Z]+"
bird_data_common_name_clean <- bird_data_common_name_clean %>% 
  mutate(species_common_name = str_replace(species_common_name, patern, ""))

```

Cleaning the species_scientific_name from the column.

```{r}
bird_data_common_name_clean <- bird_data_common_name_clean %>% 
  separate(species_scientific_name, "species_scientific_name", "\\/", remove = TRUE) 
```

Geting rid of CAPITAL LETTERS
```{r}

bird_clean <- bird_data_common_name_clean %>% 
  mutate(species_scientific_name = str_remove_all(species_scientific_name, patern))


```

Extracting column Identified
```{r}
c_bird <- bird_clean

bird_clean <- bird_clean %>% 
  mutate(Identified	 = if_else(str_detect(species_common_name,"[Uu]nidentified"),
                                          FALSE,
                                          TRUE)) %>% 
  mutate(species_common_name =str_remove(species_common_name,"\\([Uu]nidentified\\)"))
```

Deleting all rows where species_common_name =="[ BIRDS RECORDED]".
Reason: no observation noted for those records.  
```{r}
c_bird <- bird_clean
bird_clean <- c_bird %>% 
  filter(!species_common_name =="[ BIRDS RECORDED]")
```

Deleting age data from species_abbreviation column.
```{r}
c_bird <- bird_clean
bird_clean <- c_bird %>% 
  separate(species_abbreviation,into = "species_abbreviation", sep = " " )
```
Observation of the birds once considered same species. Split in 2004(8) so it stays.
```{r}
bird_clean %>% 
  filter(species_scientific_name == "Thalassarche cauta ")
```

Deleting rows with Seabird observations.
Reason: only two entries and total of 4 birds, no details also.
```{r}
c_bird <- bird_clean
bird_clean <- c_bird %>% 
  filter(species_abbreviation != "SEABUN")

```
##########Test##########Test##########Test##########Test##########Test##########Test##########Test
Outliers
```{r}
c_bird %>% 
  select(species_common_name, count) %>% 
  arrange(desc(count))
```




##########Test##########Test##########Test##########Test##########Test##########Test##########Test
Join two tables on record_id.
```{r}
full_data_seabirds <- inner_join(bird_clean,ship_data_clean_names,"record_id")

```

Selecting needed columns
```{r}
full_data_seabirds_selection <- full_data_seabirds %>% 
  select(species_common_name,
         species_scientific_name,
         species_abbreviation,
         lat,
         long,
         date,
         count)
```
##########Test##########Test##########Test##########Test##########Test##########Test##########Test

Short-tailed shearwater	
Reports of flocks size of 30 000 40 000 in 2019
Size of population deacrease by 15% evry year since 2010. 
Therefore inclination is that data collected in 70`s even though may look ext might be  somehow acurate.

Sooty shearwater
http://www.strangehistory.org/cms/index.php/archive21/86-item-10
Fairy prion	
The largest population is on Stephens Island, with 1.83 million pairs estimated. 
Prion 
Flocks can be as big as milion birds. 
```{r}
full_data_seabirds_selection %>% 
  select(species_common_name, count,lat,long,date) %>% 
  arrange(desc(count))
```



##########Test##########Test##########Test##########Test##########Test##########Test##########Test


Writing csv file
```{r}
write_csv(full_data_seabirds_selection, here("clean_data/clean_data.csv"))

```
 

