
# Homework Info

This homework recaps what you learnt in class today. 

You'll be looking at the `backpack` dataset contained within the `CodeClanData` package using tidyverse. You'll need to load the appropriate packages first.     

**In all plots you make, ensure your labels are appropriate. You also must write down a short summary of what each plot tells you in order to meet MVP.**  

**Make sure you create a new project in which to do this homework!**


# MVP

<br>

**Question 1.**  

Load in the libraries you need, and look at the `backpack` data. Take note of variable names and dimensions.   
```{r}
library(tidyverse)
library(CodeClanData)
```
```{r}
head(backpack)
```
```{r}
names(backpack)
dim(backpack)
```


**Question 2.**

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution. 
```{r}
backpack %>% 
  ggplot() +
  aes(backpack_weight)+
  geom_histogram(color = "black", fill = "green") +
  labs(
    x ="Backpack weight",
    y = "Count"
  )
```


**Question 3.**  

Are there differences in backpack weight between sex?   
```{r}
backpack %>% 
  ggplot() +
  aes(backpack_weight, fill = sex)+
  geom_histogram(alpha=0.5, color = "black" ) +
  labs(
    x ="Backpack weight",
    y = "Count"
  )
```



**Question 4.**  

What is the relationship between body weight and backpack weight?
```{r}
backpack %>% 
  ggplot()+
  aes(body_weight,backpack_weight,size =  backpack_weight) +
  geom_point(color = "black",alpha = 0.5) + 
  labs(
    x = "Body weight",
    y = " Backpack weight"
  )
```



**Question 5.**  

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.  
```{r}
backpack %>% 
  ggplot()+
  aes(body_weight,
      backpack_weight,
      shape= sex,
      color = sex) +
  geom_point(alpha = 0.9, size = 3) + 
  labs(
    x = "Body weight",
    y = " Backpack weight",
    title = "Body weight to backpack weight",
    fill = "Sex"
  )
```


**Question 6.**  

What **proportion** of students report back pain?   
```{r}
head(backpack)
sum(backpack$back_problems)
```

```{r}
backpack %>%
  ggplot()+
  aes(back_problems,fill = back_problems ==1)+
  geom_bar()+
  labs(
    x = "Percent of students with back problems",
    y = "%",
    fill = "Have a back problem"
  )
```


**Question 7.**  

Is this proportion different between men and women? 
```{r}
backpack %>%
  ggplot()+
  aes(back_problems,fill = back_problems ==1)+
  geom_bar()+
  labs(
    x = "Percent of students with back problems",
    y = "%",
    fill = "Have a back problem"
  ) +
  facet_wrap(~sex)
  
```


**Question 8.**  

Now you've investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.
```{r}
backpack %>%
  ggplot()+
  aes(back_problems,fill = back_problems ==1)+
  geom_bar()+
  labs(
    x = "Percent of students with back problems",
    y = "%",
    fill = "Have a back problem"
  ) +
  facet_wrap(~sex)
```

Try both `backpack_weight` and `body_weight` as well as the `ratio`.   

*Hint: faceting would be useful here*.   

# Extensions

<br>

**Question 1.** 

Does reported backpain differ across the year groups? Ignore year 0 in your data.  


**Question 2.** 

Is there a difference between how heavy the bags those in different years carry?


